---
title: "Covid91 vaccine study Final2020"
author: "Kelly Patrizio!"
date: "October 22,2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub-populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

$H_0$ The Null Hypothesis is that among the target population in each following data, the vaccine has no effect on whether one contracts the virus or not.

$H_A$ The Alternative Hypothesis is that among the target population in each of the following sets of data, the vaccine does have an effect on whether one contracts the virus or not.  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example sub-setting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Every question is of a Cat~Cat type.  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->  

# Plot Barplots for Gals

```{r}
DataF=filter(FinalData,sex=="gal")
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent", main="gals")
```

### Numerical Summaries For Gals

```{r}
table1 <- xtabs(~infected + treatment,data=DataF)
rowPerc(table1)
colPerc(table1)
```

### Inferential For Gals

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the frequency and percentage bar charts above for gals, it appears that the females by count who contracted COVID and took the vaccine were less likely to be infected with the virus, than those females who did not take the vaccine.  Less than 35 percent of females who took the vaccine were infected while twice as many females who didn't take the vaccine did contract the virus at approx 66 percent. It appears that the drug was effective for the gals.  Those who did not have the virus appear to show no significant effect whether they took the drug or didn't take the drug. 

According to the x-tab table in rows above for females, out of those who were infected with COVID, 33.1 percent took the vaccine and 66.9 percent didn't.  Out of those who were not infected, there appears to be no significant difference in whether they took the vaccine or not.

When we observe by count in the columns chart, it appears out of 100 tested 4.4 percent who took the vaccine were infected which is less than half of those who took the Placebo and were 8.7 percent infected.

Lastly when we observe the odds of getting the virus in the Fisher Exact Test as a female, it it appears to be a less than 50 percent chance of getting the virus if you take the vaccine.  It also appears that 95 percent of the times that we run the Fisher Exact test we will get a result in the range of 0.5 to 0.6 percent chance of getting the virus. So, overall it appears that the drug is effective for the gals.

# Plot Barplots For Guys

```{r}
DataM=filter(FinalData,sex=="guy")
barchartGC(~infected + treatment,data=DataM)
barchartGC(~infected + treatment,data=DataM, type="percent", main="guys")
```

### Numerical Summaries For Guys

```{r}
table1 <- xtabs(~infected + treatment,data=DataM)
rowPerc(table1)
colPerc(table1)
```

### Inferential For Guys

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the above frequency and percentage bar charts for guys, it appears that the guys who had been infected with COVID and took the drug were effected more than those males who did not take the drug. We also observe that those males who did not have the virus who did or did not take the drug were almost equally non-effected.  

Additionally, according to the x-tabs chart above, the men who had contracted the virus after they took the vaccine were 58.2 percent
more likely to be infected than the 41.8 percent of those who didn't and contracted the virus... Of the 3.7 percent who took the vaccine and contracted the virus, 2.6 percent didn't take the vaccine and contracted the virus. Of those who didnt contract the virus, there was no significant difference whether the took the vaccine or not! Overall, men who took the vaccine were infected more than those who didn't and it would appear that the vaccine was not effective for the guys.

When we observe the Fisher Exact Test above it appears that the odds of getting the virus if you take the vaccine is probable at a ratio of 1.5 percent.  As many times as the test is run 95 percent of the time the results would run from 1.2 to 1.8 ratio.

# Plot Barplots For LGBTQ

```{r}
DataH=filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataH)
barchartGC(~infected + treatment,data=DataH, type="percent", main="gay")
```


### Numerical Summaries For LGBTQ

```{r}
table1 <- xtabs(~infected + treatment,data=DataH)
rowPerc(table1)
colPerc(table1)
```

### Inferential For LGBTQ


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the above frequency and percentage bar charts, it appears that 71 percent of those who were gay and got the vaccine, contracted the virus, making them more likely to be infected than those who didn't take the drug and at a much lower rate of 30 percent.  Those who didn't contract the virus appear to have no significant difference whether taking the drug or not.

According to the x-tabs chart above among the LGBQT community 70.7 percent who contracted the virus took the vaccine and 29.3 percent didn't.  Those from our count data just under 8 percent who took the vaccine were infected by the virus, the other 3.4 percent didnt take the vaccine and also contracted the virus.  it appears that twice as many who took the vaccine contracted the virus than those who didnt take the vaccine.  it appears the vaccine is not effective among gays. 
   
As we take a look at the Fisher Exact Test, we can see similar results with a probability ratio of 2.5 percent of gays contracting the virus.  as many times that this teswt is conduct6ed 95 percent of the time there is anywhere from 1.6 to 3.9 percent chance of contracting the virus if your gay and take the vaccine. 

# Plot Barplots For Drug Users

```{r}
DataU=filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=DataU)
barchartGC(~infected + treatment,data=DataU, type="percent", main="Yes")
```

### Numerical Summaries For Drug Users

```{r}
table1 <- xtabs(~infected + treatment,data=DataU)
rowPerc(table1)
colPerc(table1)
```

### Inferential For Drug Users

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the frequency and percentage bar charts above, it appears that those who were drug users and did receive the vaccine, were less effected than those who didn't receive the drug. At 33.3 percent who contracted the virus, took the vaccine; 66.7 percent didn't. Those who didn't contract the virus, whether they took the drug or not, showed no significant difference.  

Similarly, according to the x-tabs table above, out of those in our data, 33.3 percent who contracted the virus took the vaccine, and 66.7 didn't. The 3.6 percent in our count data who took the vaccine contracted the virus and 6.5 percent who didn't take the vaccine also contracted the virus. It appears that the vaccine is effective for drug users since twice as many drug users who didn't get the vaccine did contract the virus.

The Fisher Exact Test appears to show that the vaccine is effective among drug users at a probability of getting the virus ratio of 0.5 chance.  as many times as this test is conducted, 95 percent of the time a ratio of 0.4 to 0.8 will contract the virus.   
 
 



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->  

In conclusion, we reject the NULL Hypothesis that among the target population for "gals" and "drug users" in the above sets of data, that the vaccine has no effect on whether one contracts the virus or not, in favor of the ALTERNATIVE hypothesis that among the target population for "gals" and "drug users" in the above sets of data, the vaccine does have an effect on whether one contracts the virus or not. 

In the case of the "guys" and the "gays" we fail to reject the NULL hypothesis for lack of significant evidence that the vaccine would have a positive effect!

It appears that between genders, women were less likely to contract the virus when taking the vaccine than men. 

It appears that gays were more likely to contract the virus than straights when taking the vaccine.  

Lastly, it appears that drug users were less likely to contract the virus than non-drug users when taking the vaccine.

Overall, we can conclude that among the sub-populations whether taking the vaccine or not taking the vaccine can have a negative or positive effect in relation to their sub-groups such as the above sub-groups between genders, and various lifestyles.