---
title: "Covid91 vaccine study Final2020"
author: "Kelly Patrizio!"
date: "October 22,2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub-populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

$H_0$ The Null Hypothesis is that among the target population in each following data, the vaccine has no effect on whether one contracts the virus or not.

$H_A$ The Alternative Hypothesis is that among the target population in each of the following sets of data, the vaccine does have an effect on whether one contracts the virus or not.  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example sub-setting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Every question is of a Cat~Cat type.  

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->  

# Plot barplots for Gals


```{r}
DataF=filter(FinalData,sex=="gal")
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent", main="gals")

```

In the frequency bar chart above, titled gals, it appears that the females who had COVID those who took the drug were less effected with the virus, than those females who did not take the drug.  Those who did not have COVID appear to not be effected whether they took the drugh or didn't take the drug. 

As we observe the percentage bar chart above, titled gals, it appears that similar to the frequency chart, that those females who had the virus, the females who took the drug were less effected than those who didn't take the drug and those females who didn't have the drug.

### Numerical Summaries For Gals

```{r}
table1 <- xtabs(~infected + treatment,data=DataF)
rowPerc(table1)
colPerc(table1)
```

According to the Cross tab table in rows above for females, out of those who were infected with COVID, 33% took the drug and 67% didn't.  Out of those who were not infected, there appears to be no significant difference in whether they took the vaccine or not. 

### Inferential For Gals


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

# Plot Barplots For Guys


```{r}
DataM=filter(FinalData,sex=="guy")
barchartGC(~infected + treatment,data=DataM)
barchartGC(~infected + treatment,data=DataM, type="percent", main="guys")

```

### Numerical Summaries For Guys

```{r}
table1 <- xtabs(~infected + treatment,data=DataM)
rowPerc(table1)
colPerc(table1)
```

### Inferential For Guys


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


# Plot Barplots For LGBTQ

```{r}
DataH=filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataH)
barchartGC(~infected + treatment,data=DataH, type="percent", main="gay")
```


### Numerical Summaries For LGBTQ

```{r}
table1 <- xtabs(~infected + treatment,data=DataH)
rowPerc(table1)
colPerc(table1)
```

### Inferential For LGBTQ


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```




In the above frequency bar chart, for guys, it appears that the males who had COVID and took the drug were effected more than those males who did not take the drug. We also observe that those males who did not have the virus who did or did not take the drug were equally almost equally non-effected.  

Likewise when we observe the bar chart by percentage above for guys, it appears that those males who had the virus and took the drug were more effected with the Virus than those who didn't take the drug, and those who didn't have the virus, were almost equally non-effected whether they took the drug or not. 




In the frequency bar chart above it appears that those who were straight and had COVID, and took the drug were less effected than those who didn't take the drug.  Those who did not have the virus, it appears that there wasn't a significant effect on whether or not they took the drug or not. 

Similarly, when we observe the bar chart in percentages, those who were straight and had the virus and took the drug were less likely to get the virus than those who did take the drug.  Those who didn't have the virus, whether they took the drug or not, did not show a significant effect either way. 

```{r}
DataH=filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataH)
barchartGC(~infected + treatment,data=DataH, type="percent", main="gay")

```

In the above frequency bar chart, we observe that those who were gay and got the virus, were more likely to be effected when taking the drug than those who didn't take the drug.  Those who didn't contract the virus and were gay, appear to have no significant effect whether taking the drug or not.

Similarly, in the percentage bar chart, those who were gay and took the drug were significantly more effected with the virus when taking the drug than those who didn't take the drug, and those who were not effected showed no significant difference whether taking the drug or not.  

# Plot Barplots For Drug Users

```{r}
DataU=filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=DataU)
barchartGC(~infected + treatment,data=DataU, type="percent", main="Yes")
```

### Numerical Summaries For Drug Users

```{r}
table1 <- xtabs(~infected + treatment,data=DataU)
rowPerc(table1)
colPerc(table1)
```

### Inferential For Drug Users


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the frequency bar chart above, we observe that those who were drug users and did receive the vaccine, appear to be less effected than those who didn't receive the drug, and those who didn't contract the virus, whether they took the drug or not, showed no significant difference.  

Similarly, when we observe the percentage bar chart above, it appears that those drug users who contracted the virus and took the vaccine were less likely to contract the virus than those who didn't take the vaccine.  Those who didn't contract the virus showed no apparent difference between those who took the vaccine and those who didn't and were drug users. 



  




According to the Cross tab table in columns above for females, out of those who took the drug 4% contracted the virus and 96% didn't.  Out of those who didn't take the drug, 9% contracted the virus and 91% didn't.








```{r}
DataH=filter(FinalData,LGBTQ=="straight")
barchartGC(~infected + treatment,data=DataH)
barchartGC(~infected + treatment,data=DataH, type="percent", main="straight")

```



According to the xtabs chart above the men who had contracted the virus after they took the vaccine were more likely to be infected than those who didn't and of those who didn't contract the virus and took the vaccine, were less likely to be infected by the virus than those who didn't take the vaccine.  When the study was completed, 3.72 percent of men who took the vaccine were infected by the virus and 96.28 percent who did take the vaccine were not infected by the virus. Overall, men who took the vaccine were infected more than those who didn't.


```{r}
table3 <- xtabs(~infected + treatment,data=DataH)
rowPerc(table3)
colPerc(table3)
```

In the xtabs chart above among the LGBQT community 70.69 percent of those who were infected by the virus, took the vaccine; a substantial difference to those who didn't take the vaccine.

Those in the LGBQT community who took the vaccine 7.97 percent were infected with the virus and 3.36 percent were not infected that didn't take the vaccine.  Overall the vaccine was not as effective as those who didn't take the vaccine.    

```{r}
table1 <- xtabs(~infected + treatment,data=DataU)
rowPerc(table1)
colPerc(table1)
```

According to the xtabs chart above, those who are drug users that were infected by the virus 43.29 percent took the vaccine and 56.71 percent who did not take the vaccine, were infected. Out of 4.1 percent of drug users who took the vaccine and were infected, 95.9 percent who also took the vaccine were not infected. Overall, drug users infected were less likely to be infected than those who didnt take the vaccine.  However, it appears that there is not a significant difference between those who took the vaccine and those who didn't.  



The above Chi-Square Statistic Chart shows an overall P-value of approx. .000000000005 which is below the 0.05 and true odds ratio is not equal to 1 making it impossible to prove a non relation to the targeted sets of data ran above.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the George Town videos collection.!-->  

In conclusion, we reject the null Hypothesis that among the target population in each following data, the vaccine has no effect on whether one contracts the virus or not in favor of the alternative hypothesis that among the target population in each of the following sets of data, the vaccine does have an effect on whether one contracts the virus or not.  

It appears that between genders, women were less likely to contract the virus when taking the virus than men. 

It appears that gays were more likely to contract the virus than straights when taking the vaccine.  

Lastly it appears that drug users were more likely to contract the virus than non-drug users when taking the vaccine.

Overall, we can conclude that among the sub-populations whether taking the vaccine or not taking the vaccine can have a negative or positive effect in relation to their sub-groups such as the above sub-groups between genders, and various lifestyles.